---
title: "An Investigation On Gold Price and Inflation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Gold Price and Inflation

Today, I come across an article about the relationship between gold price and inflation. There is an arguement in the article stating that "inflation is a bad thing for gold".
The logic goes like "if the inflation is high, then securities like bonds will be required a higher yield. Since gold has no yield, investors will sell gold and buy bonds for higher yield. Therefore, gold price goes down.”

Still, I feel somewhat unconvinced by this logic. 

Well, why not let data speak itself. 

First, i got the gold price data and inflation data from Quandl API

```{r }
library(Quandl)#financial data
library(data.table)#read table
library(lubridate)#date manupulation
library(scales)
library(ggthemes)#change plot themes
library(dplyr)#dataframe
library(ggplot2)
gold <- Quandl("LBMA/GOLD")#gold data
api <- "https://www.quandl.com/api/v3/datasets/FRED/DTB3.csv?api_key=prFMytag1uYoMRJLGyNa"
inflation = fread(api)#inflation data
# get year, month , day off the Date
gold$Date <- as.Date(gold$Date, "%y/%m/%d")
gold$year <- year(ymd(gold$Date))
gold$month <-  month(ymd(gold$Date))
gold$day <- day(ymd(gold$Date))
inflation$year <- year(ymd(inflation$Date))
inflation$month <-  month(ymd(inflation$Date))
inflation$day <- day(ymd(inflation$Date))

```

with data in hand, let's first plot some simple chart

first, I ploted the Inflation over time, as you may have noticed, inflation peaked  during 1980s, complex period
```{r}
P1 <- ggplot(inflation, aes(x = factor(year),y = Value))+
  geom_boxplot()+
  scale_x_discrete(name = 'Year',breaks=seq(1950, 2020, 10))
  
P1 + theme_economist() + scale_colour_economist() +
  scale_y_continuous(name= "Inflation Rate", position = "left")+
  ggtitle("Inflation Over Time")
```

then, i'll plot the Gold Price Over Time
```{r}
gold$USD <- gold[2] 

P2 <- ggplot(gold, aes(x = factor(year),y = USD ))+
  geom_boxplot()+
  scale_x_discrete(name = 'Year',breaks=seq(1950, 2020, 10))
P2 + theme_economist() + scale_colour_economist() +
  scale_y_continuous(name= "Gold Price", position = "left")+
  ggtitle("Gold Price Over Time")
```

next, let's figure out the relationship between gold price and inflation. As you can see ,they have a quite strong negtive relationship , aren't they? 
However, the inflation is relative figure, we may have to turn gold price the same way as inflation. 

```{r }

inf_y <- select(inflation,year,Value)# yearly data for inflation
ys <- summarise(group_by(tbl_df(inf_y),year),mean(Value))#yearly data for inflation

gold$USD <- gold$`USD (AM)`

gs <- select(gold,year,USD)# yearly data for gold

gsd <- summarise(group_by(tbl_df(gs),year),mean(USD))# yearly data for gold

com = merge(x= ys,y = gsd,by="year",all= TRUE)#yearly data for gold and inflation
com$inflation <- com$`mean(Value)`#change name
com$USD <- com$`mean(USD)`#change name

com <- select(com, year, USD, inflation)#drop value
com <- na.omit(com)#get rid of na
com$USD <- (com$USD-mean(com$USD))/sd(com$USD)#norm and std
com$inflation <- (com$inflation-mean(com$inflation))/sd(com$inflation)#norm and std
cor(com$USD,com$inflation)#correlation



```

next, let's see their relationship again 

```{r}
lg <- filter(com, year>1968)#from 2ed row
sm <- filter(com, year < 2017)#to -1 row
ts <- filter(com,year>1968)#third data

ts$rtg <- lg$USD/sm$USD#relative gold price
ts$inflation <- lg$inflation#inflation
tss <- ts[c('year','rtg','inflation')]#year ,relative gold price , inflation
cor(tss$rtg,tss$inflation)
```


last, let's draw the picture to visualize it. 

```{r}
tsss <- melt(tss,id='year')#prepare for drawing
P3 <- ggplot(tsss,aes(x = year,y = value,colour=variable,group=variable,alpha=0.8,size=2))+
  geom_line()+#usd line
  scale_x_continuous(name = 'Year',breaks=seq(1950, 2020, 10))
P3 + theme_economist() + scale_colour_economist() +
  scale_y_continuous(name= "standard and Normaized price", position = "left")+
  ggtitle("Gold Price with Infation")
```